---
title: "Evolutionary Dynamics"
author: xinyu yang
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pre-step1: load functions

## Basic function
- `load_ccf(sample,input)`
- `file_format(filename,samplenamecol`
## Step1: Ccube input format 
`ParseSnvCnaPcawgFormat(ssm,cna)`
## Step2: Build post summary for ccube samples
`Build_post_summary(input,output=NA)`
-input : Input Folders which stored Ccube results (Ccube_folder)
-output : output folder to store the post summary file, only return `post_summary` if `output=NA`
## Step3: Construct ccf matrix
`CountMatBuild(samplelist,genelist=NA,rows=101,input_folder)` | Build CCF Matrix for chosen samples with bin width 0.01
- rows: specicfy rows, eg. if we choose ccf band 0-1, set rows=101
- upper: specify ccf upper limit
- input_folder: folder stored CCF files (Ccube_folder)
`TypeMatrixBuild(post_summary,input_folder,output,ccfupper=1,RankEstimateType="fraction")`
- post_summary : obstained from step 1
- input_folder : Input Folders which stored Ccube results (Ccube_folder)
- output : output folder to store CCF matrix (Matrix_folder)
- ccfupper: specify ccf upper limit 
- RankEstimateType : output file type to do further rank estimate analysis
## Step 4: Rank estimate (python) + plot(R)
`rank_estimate_plot(outputFolder)`: plot rank estimate csv in outputFolder and save the plot there
## Step 5: NMF for Each type
`sig_plot` : plot sig matrix
`nmf_sig_plot_type`|`sig_plot`: Run NMF for specified cancertype with specified rank 
- type: specify cancertype
- input_folder: folder stored CCF matrix (Matrix_folder)
- output : folder to store NMF results (NMF_folder)
- rank_summary: summary of rank estimate results (visual inspection)
`nmf_sig_all_plot`|`nmf_sig_plot_type`: output all sig plot for all cancertype
- input_folder: Folders stored Matrix ('/CountMatrix/2019-12-28/')
- output: Folders stored nmf results ('/NMF_results/2019-12-28/')
- rank_summary: summary of rank estimate results (visual inspection,location "/rank_summary.csv")
## Step 6: Combine sig and run Hierarchical cluster
`combine_nmf(input_folder,cancertype,return="sig"/"expo")`: output combined sig/expo matrix
- input_folder: folders stores NMF result (NMF_folder)
## Step 7: test the number of cluster
## Step 8: run Hierarchical clustering and plot (save to 8.HC_consensus)
```{r}
file_format <- function(filename=filename,samplenamecol){
  names <- colnames(filename)
  names[samplenamecol] <- "samplename"
  names -> colnames(filename)
  filename$samplename <- substr(filename$samplename,1,12)
  filename$samplename <- gsub("[.]","-",filename$samplename)
  return(filename)
}

load_ccf <- function(samplename,input){
  Check <- ArgumentCheck::newArgCheck()
  suppressWarnings(rm(ssm,res,ccubeRes))
  
  format1 <- paste0(input,samplename,"/ccube_result.RData")
  format2 <- paste0(input,samplename,"/ccube_res_v0.3_final_new.RData")
  format3 <- paste0(input,samplename,"/ccube_res_run1.RData")
  format4 <- paste0(input,samplename,"/ccubeRes.RData")
  
  if (file.exists(format4 )) load(format4)  
    else if (file.exists(format1 )) load(format1)  
     else if(file.exists(format2)) load(format2) 
      else if(file.exists(format3)) load(format3) 
        else{
          ArgumentCheck::addError(
          msg = "No file has been loaded",
          argcheck = Check)
          }
  
  if (exists("ccubeRes")) return(ccubeRes$ssm)  else
     if (exists("res")) return(res$ssm) else
       if (exists("ssm")) return(ssm)
}

## Step1: Ccube input format - ParseSnvCnaPcawgFormat(ssm,cna)
ParseSnvCnaPcawgFormat <- function (ssm, cna) {
      library(dplyr)
      
      ssm <- ssm %>%
        mutate(chr= substr(chr,4,length(chr)),
               cn_frac = NA,
               major_cn = NA,
               minor_cn = NA,
               mutation_id = NA)
      
      for (jj in seq_len(nrow(cna)) ) {
        cc = cna[jj,]
        
        idx = which(ssm$chr == cc$chromosome &  (ssm$Start_Position >= cc$start & ssm$End_Position <= cc$end) )
        
        if (length(idx) > 0) {
          ssm[idx,] <- ssm[idx,] %>% 
            mutate( major_cn=cc$major_cn,minor_cn =cc$minor_cn, cn_frac = 1)
        }
      }
      
      ssm$mutation_id = paste0(ssm$chr, "_", ssm$Start_Position )
      
      ssm <- ssm %>%
        select(-chr,-Start_Position,-End_Position,-df.n_alt_count,-n_ref_count) %>%
        rename(var_counts='t_alt_count',ref_counts='t_ref_count') %>%
        mutate(total_counts=var_counts+ref_counts,normal_cn=2) %>%
        filter(!is.na(major_cn) ,!is.na(minor_cn),!is.na(cn_frac),major_cn > 0)
      
      return(ssm)
}


Build_post_summary <- function(input,output=NA,mergeType= F, typefile="NA"){
  library(magrittr)
  library(dplyr)
  
  sample_list <- dir(input)
  
  colnames <- c("samplename","Tumor_Sample_Barcode","n_mutations","ccf_0-1_percentage","ccf_0-2_percentage","Ncluster","purity","ccube_purity","ccf_mean_cluster1","ccf_mean_cluster2","ccf_mean_cluster3","ccf_mean_cluster4","ccf_mean_cluster5")

  # if (i==1) cat("Start building post summary for",length(sample_list),"files","\n")
  # if (i%%1000==0) print(paste0("----- Finish loading ",i," files -----"))

  post_summary_analyse <- function(samplename){
      library(dplyr)
      suppressWarnings(rm(ssm))
      ssm <- load_ccf(samplename,input=input)
      ccf <- unique(ssm$ccube_ccf_mean)
      ccf_mean_order <- sort(ccf,decreasing = T)
      Ncluster <- length(ccf)
      
      post_summary <- data.frame(samplename <- samplename) %>%
        mutate(
          Tumor_Sample_Barcode = unique(ssm$Tumor_Sample_Barcode),
          n_mutations = nrow(ssm),
          ccf_01_percentage = mean(ssm$ccube_ccf<=1),
          ccf_02_percentage = mean(ssm$ccube_ccf<=2),
          Ncluster = Ncluster,
          purity = unique(ssm$purity),
          ccube_purity = ifelse(exists("ssm$ccube_purity"),unique(ssm$ccube_purity),NA),
          ccf_mean_cluster1 = ifelse(Ncluster>=1,ccf_mean_order[1],0),
          ccf_mean_cluster2 = ifelse(Ncluster>=2,ccf_mean_order[2],0),
          ccf_mean_cluster3 = ifelse(Ncluster>=3,ccf_mean_order[3],0),
          ccf_mean_cluster4 = ifelse(Ncluster>=4,ccf_mean_order[4],0),
          ccf_mean_cluster5 = ifelse(Ncluster>=5,ccf_mean_order[5],0)
        ) %>% set_colnames(colnames)
      return(post_summary)
     }
     
  post_summary <- do.call(rbind,lapply(sample_list,post_summary_analyse))
  
  if (mergeType==T & !is.na(typefile)) {
    cancertype <- read.csv(file=typefile)[,-1] 
    colnames(cancertype)[1] <- "samplename"
    post_summary <- left_join(post_summary, cancertype,by="samplename") 
    if ("Types" %in% colnames(post_summary)) colnames(post_summary)[which(colnames(post_summary)=="Types")] <- "cancertype"
  }
  
  if (!is.na(output)) {
    write.csv(post_summary,file=paste0(output,"post_summary_",length(sample_list),"_",Sys.Date(),".csv"))
    save(post_summary,file=paste0(output,"post_summary_",length(sample_list),"_",Sys.Date(),".RData"))
  }
  return(post_summary)
}

# Build CCF Matrix for chosen samples within region 0-upper
CountMatBuild <- function(samplelist,upper,input_folder,genelist=NA){
  library(dplyr)
  
  n_sample <- length(samplelist)
  rows <- upper/0.01 + 1
  
  ccfBand <- seq(0,upper,length.out = rows)
  ccfBandCountsMat <- matrix(nrow=rows,ncol=n_sample)
  
  if (n_sample ==0) stop("The number of choosen sample is 0")
  
  for (j in 1:n_sample){
    sample_name <- as.character(samplelist[j])
    ssm <- load_ccf(sample_name,input=input_folder)
    if (!is.na(genelist)) ssm <- subset(ssm,SYMBOL %in% genelist)
    matchBandCountDf <- data.frame(Var1=as.character(1:rows))
    matchBandCountDf <- suppressWarnings(left_join(matchBandCountDf,as.data.frame(table(findInterval(ssm$ccube_ccf,ccfBand))),stringAsFactors = F,by="Var1"))
    ccfBandCountsMat[,j] <- matchBandCountDf$Freq
  }
  # delete the final row
  ccfBandCountsMat <- ccfBandCountsMat[-rows,]
  ccfBandCountsMat[is.na(ccfBandCountsMat)] <- 0
  
  if (n_sample==1) {ccfBandFractionMat <- ccfBandCountsMat/sum(ccfBandCountsMat)
                    ccfBandFractionMat.random <- ccfBandFractionMat[sample(1:(rows-1))]
                    ccfBandCountsMat.random <- ccfBandCountsMat[sample(1:(rows-1))]} else 
    {ccfBandFractionMat <- apply(ccfBandCountsMat,2,function(x) x/sum(x))
     ccfBandCountsMat.random <- NMF::randomize(ccfBandCountsMat)
     ccfBandFractionMat.random <- apply(ccfBandCountsMat.random,2,function(x) x/sum(x))
     }
    
  return(list(ccfBandCountsMat,ccfBandCountsMat.random,ccfBandFractionMat,ccfBandFractionMat.random))
}

# Build CCF Matrix for all cancer type with >30 samples 


TypeMatrixBuild <- function(post_summary,input_folder,output,ccfupper=1,RankEstimateType="fraction"){
  
  type <- post_summary %>%
    group_by(cancertype)%>%
    summarize(n=n())
  
  typelist <- as.character(subset(type,n>=30)$cancertype)
  
  ntype <- length(typelist)
  
  print(paste0("Construct ccf count matrix for ",ntype," types (> 30 samples)"))

  if (!dir.exists(output)) dir.create(output)
  
  for (i in 1:ntype){
    
    library(NMF)
    library(dplyr)
    
    # get sample list for each type
    type <-  typelist[i]
    
    print(paste0("finish load ",i,'th type - ',type))
    
    samplelist <- subset(post_summary,cancertype==type)$samplename
    n_sample <- length(samplelist)
    
    ccfBandCountsMat <- suppressWarnings(CountMatBuild(samplelist,input_folder = input_folder,upper=ccfupper))
 
    if (!is.na(output)){
        
        ccfCountMatrix <-ccfBandCountsMat[[1]]
        ccfCountsMatrix.random <- ccfBandCountsMat[[2]]
        ccfFractionMatrix <- ccfBandCountsMat[[3]]
        ccfFractionMatrix.random <- ccfBandCountsMat[[4]]
        
        # create direction
        if (!dir.exists(paste0(output,type,"/"))) dir.create(paste0(output,type,"/"))  
        if (!dir.exists(paste0(output,"rank_estimate/"))) dir.create(paste0(output,"rank_estimate/"))  
                                                                    
        # Output Matrix 
        output_format <- paste0(output,type,"/",type,"_", n_sample,"_0-",ccfupper)
        output_rank <- paste0(output,"rank_estimate/",type,"_", n_sample,"_0-",ccfupper)
      
        write.csv(samplelist,file=paste0(output,type,"/",type,"_",n_sample,"_samplelist_",Sys.Date(),".csv"))
        
        save(ccfCountMatrix,file=paste0(output_format,"_ccfCountMatrix_",Sys.Date(),".RData"))
        save(ccfCountsMatrix.random,file=paste0(output_format,"_ccfCountsMatrix.random_",Sys.Date(),".RData"))
        save(ccfFractionMatrix,file=paste0(output_format,"_ccfFractionMatrix_",Sys.Date(),".RData"))
        save(ccfFractionMatrix.random,file=paste0(output_format,"_ccfFractionMatrix.random_",Sys.Date(),".RData"))
        
        # output to rank estimate folder
        if (RankEstimateType=="fraction") {
          write.csv(ccfFractionMatrix.random,file=paste0(output_rank,"_ccfFractionMatrix.random_",Sys.Date(),".csv"))
          write.csv(ccfFractionMatrix,file=paste0(output_rank,"_ccfFractionMatrix_",Sys.Date(),".csv"))
        }
        
        if (RankEstimateType=="count") {
          write.csv(ccfCountsMatrix.random,file=paste0(output_rank,"_ccfCountMatrix.random_",Sys.Date(),".csv"))
          write.csv(ccfCountMatrix,file=paste0(output_format,"_ccfCountMatrix_",Sys.Date(),".csv"))
        }
    }
  }
}


rank_estimate_plot <- function(outputFolder,rankfilepath) {
  library(ggplot2)
  library(ggpubr)
  library(gridExtra)
  library(dplyr)
  file <- unique(unlist(lapply(dir(outputFolder),function(x) strsplit(x,"_")[[1]][[1]])))
  rank_blank <- data.frame(cancertype=file,rank=NA,rss_suggested_rank=NA)
  

  for (i in 1:length(file)) {
    tryCatch({
        filename <- file[i]
        estimate <- read.csv(paste0(outputFolder ,filename,"_ccfFractionMatrix.csv")) %>% mutate(type='normal')
        estimate_random <- read.csv(paste0(outputFolder ,filename,"_ccfFractionMatrix.random.csv")) %>% mutate(type='random')
        
        print(paste0("plot for ",filename)) 
        estimate_rank <- rbind(estimate,estimate_random)
        
        rss_decrease <-  which((estimate[order(estimate$rank),]$rss[1:10]-estimate[order(estimate$rank),]$rss[2:11]) - (estimate_random[order(estimate_random$rank),]$rss[1:10]-estimate_random[order(estimate_random$rank),]$rss[2:11])<0)[1]
        rank_blank[which(rank_blank$cancertype == filename),]$rss_suggested_rank <- rss_decrease + 1
         
        write.csv(estimate_rank,paste0(outputFolder,filename,'_rank_summary.csv'))
        
        pdf(paste0(outputFolder,filename,'_rank_summary.pdf'),width = 12,height = 6)
        p1 <- ggplot(data=estimate_rank,aes(x=rank,y=cophenetic,group=type)) + geom_line(aes(color=type))+geom_point(aes(color=type))+theme(legend.position = "none",axis.title.x=element_blank())
        p2 <- ggplot(data=estimate_rank,aes(x=rank,y=dispersion,group=type)) + geom_line(aes(color=type))+geom_point(aes(color=type))+theme(legend.position = "none",axis.title.x=element_blank())
        p3 <- ggplot(data=estimate_rank,aes(x=rank,y=evar,group=type)) + geom_line(aes(color=type))+geom_point(aes(color=type))+theme(legend.position = "none",axis.title.x=element_blank())
        p4 <- ggplot(data=estimate_rank,aes(x=rank,y=rss,group=type)) + geom_line(aes(color=type))+geom_point(aes(color=type))+theme(legend.position = "none",axis.title.x=element_blank())
        p5 <- ggplot(data=estimate_rank,aes(x=rank,y=euclidean,group=type)) + geom_line(aes(color=type))+geom_point(aes(color=type))+theme(legend.position = "none",axis.title.x=element_blank())
        p6 <- ggplot(data=estimate_rank,aes(x=rank,y=kl,group=type)) + geom_line(aes(color=type))+geom_point(aes(color=type))+theme(legend.position = "bottom",axis.title.x=element_blank())
        p7 <- ggplot(data=estimate_rank) + geom_line(aes(x=rank,y=sparseness1,group=type,color=type))+geom_line(aes(x=rank,y=sparseness2,group=type,color=type))+geom_point(aes(x=rank,y=sparseness2,group=type,color=type))+theme(legend.position = "none",axis.title.x=element_blank())
        
        g_legend<-function(a.gplot){
          tmp <- ggplot_gtable(ggplot_build(a.gplot))
          leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
          legend <- tmp$grobs[[leg]]
          return(legend)}
        
        mylegend <-g_legend(p6)
        
        print(grid.arrange(arrangeGrob(p1,p2,p3,p4,p5,p6+theme(legend.position = "none"),nrow=2),mylegend,nrow=2,top=paste0("NMF rank estimate for ",filename),heights=c(10, 1)))
        dev.off()
    },error=function(e) print("error!"))
  }
  
  write.csv(rank_blank,file=rankfilepath)
}

sig_plot <- function(sig){
  library(gridExtra)
  library(ggplot2)
  library(RColorBrewer)
  library(grid)
  library(dplyr)
  library(reshape2)
  library(cowplot)
  library(magrittr)
  
  xx <- as.data.frame(t(apply(sig,2,function(x) x/sum(x)))) %>%
    set_colnames(1:ncol(.)) %>%
    mutate(signature=paste0("Signature ",1:nrow(.))) %>%
    melt(.,id=c("signature"))
  
  fills <- brewer.pal(8, "Set3")[c(1,3:8,2)]
  
  p1 <- ggplot(xx,aes(y=value,x=variable)) + geom_bar(aes(fill=signature),stat='identity') + scale_fill_manual(values = fills)+ theme_grey()+
    #ggtitle(paste0("rank = ",rank,", cancertype = ",type, ", MatrixType = ",MatType )) + 
    theme(legend.title = element_blank(),
          legend.position = "none",
          strip.text.x = element_text(color= "white",size=10),
          panel.grid.minor.y = element_blank(),
          axis.title.x = element_text(color = "grey20", size = 8),
          axis.text.x = element_text(color = "grey20", size = 6),
          axis.text.y = element_text(color = "grey20", size = 6),
          plot.title = element_text(size= 10)) +
    facet_grid(cols = vars(signature))+ scale_x_discrete(breaks=c("1","50","100") ,labels=c("0", "0.5", "1"))+xlab("Cancer Cell Fraction") + ylab("") 
  
  g1 <- ggplot_gtable(ggplot_build(p1))
  
  strip_both <- which(grepl('strip-', g1$layout$name))
  
  k <- 1
  for (i in strip_both) {
    j1 <- which(grepl('rect', g1$grobs[[i]]$grobs[[1]]$childrenOrder))
    g1$grobs[[i]]$grobs[[1]]$children[[j1]]$gp$fill <- fills[k]
    #g1$layout$clip[j3] <- "off"
    k <- k+1
  }
  grid.draw(g1)
  #output sig and expo
  return(g1)
}

nmf_sig_plot_type <- function(type,MatType="fraction",input_folder,output,rank_summary){
     library(magrittr)
  
     type_path <- paste0(input_folder,type,"/")
     if (MatType=="fraction"){
     file_path <- paste0(input_folder,type,"/",dir(type_path)[grep("ccfFractionMatrix_",dir( type_path))])
     } 
  
     if (MatType=="count"){
     file_path <- paste0(input_folder,type,"/",dir(type_path)[grep("ccfCountsMatrix_",dir( type_path))])
     }
     
     samplename_path <- paste0(input_folder,type,"/",dir(type_path)[grep("samplelist",dir(type_path))]) 
     
     if (!dir.exists(paste0(output,type))) dir.create(paste0(output,type))
       
     load(file=file_path)
     samplename <- as.character(read.csv(file=samplename_path)[,-1])
     n_sample <- length(samplename)
     
     #format rank summary file
     rank <- as.numeric(subset(rank_summary,cancertype == type)$rank)
     
     if (exists("ccfFractionMatrix")) ccfMat <- ccfFractionMatrix
     if (exists("ccfCountsMatrix")) ccfMat <- ccfCountsMatrix
     
     index <- sample(1:ncol(ccfMat))
     ccfMat <- ccfMat[,index]
     samplename_random <- samplename[index]
     
     #preprocess for rows with all 0
     index_p <- which(rowSums(ccfMat)>0)
     index_n <- which(!rowSums(ccfMat)>0)
     ccfMat <- ccfMat[which(rowSums(ccfMat)>0),]
     
     
     #run NMF
     res <- nmf(ccfMat,rank,.opt='vp4')
     sig <- as.data.frame(matrix(0,nrow=length(index_p)+length(index_n),ncol=ncol(res@fit@W)))
     sig[c(index_p),] <- as.data.frame(res@fit@W) %>%
       set_colnames(paste0("sig_",1:ncol(.)))
     
     expo <- as.matrix(res@fit@H)
     
     #sig plot
     
     p_sig <- sig_plot(sig)
   
     #output sig and expo
     expo <- as.data.frame(t(expo)) %>%
        cbind(.,samplename_random) 
     colnames(expo)[1:rank] <- paste0("sig_",1:rank)
     
     save(expo,file=paste0(output,type,'/',type,'_',n_sample,"_expowithsample_",Sys.Date(),".RData"))
     save(sig,file=paste0(output,type,'/',type,'_',n_sample,"_sig_",Sys.Date(),".RData"))
     save(res,file=paste0(output,type,'/',type,'_',n_sample,"_res_",Sys.Date(),".RData"))
     
     
     return(list(p_sig,n_sample))
}

nmf_sig_all_plot <- function(input_folder,output,rank_summary) {
  library(gridExtra)
  library(ggplot2)
  library(grid)
  library(dplyr)
  library(reshape2)
  library(cowplot)
  library(NMF)
  
  if (!dir.exists(output)) dir.create(output)
  
  colnames(rank_summary) <- c("cancertype","rank")
  cancertype <- as.character(rank_summary$cancertype)
 
  j = 0
  for (i in 1:length(cancertype)) {
    tryCatch({
    type <- cancertype[i]
   
    print(paste0("load ",i,"th type - ",type))
    
    nmf_result <- nmf_sig_plot_type(type,input_folder=input_folder,output=output,rank_summary=rank_summary)
    p_sig <- nmf_result[[1]]
    n_sample <- nmf_result[[2]]
  
    p <- plot_grid(p_sig)
    title <- ggdraw() + draw_label(paste0("Extracting Signatures for ",type," with ",n_sample," samples "),fontface='bold')
    j <- j+1
    commands <- paste0("p",j," <- plot_grid(title,p,align='V',ncol=1,rel_heights = c(0.1,0.15,0.65))")
    eval(parse(text=commands))
    },error = function(e) print(paste0("Failed run NMF on this type")))
  }
    #plot_grid(p1)
    
    ntype <- length(cancertype)
    
    n_file <-  j %% 8
    
    if (n_file==0) n_pdf <- (j %/% 8) else
      n_pdf <- (j %/% 8)+1
    
    for (i in 1:n_pdf){
     
      if (i != n_pdf) {
        commands1 <- paste0("pdf(file=paste0(output,'Sig_summary_p",(8*i-7),"-",8*i,"_',Sys.Date(),'.pdf'),height=25,width=12)")
        commands2 <- paste0("g",i,"<- plot_grid(",paste0("p",(8*i-7):(8*i),collapse=","),",align='V',ncol=1,rel_heights = c(0.1,0.1,0.1,0.1,0.1,0.1,0.1,0.1))") 
      } else {
        
        commands1 <- paste0("pdf(file=paste0(output,'Sig_summary_p",(8*i-8),"-",8*i-8+n_file,"_',Sys.Date(),'.pdf'),height=25,width=12)")
        commands2 <- paste0("g",i," <- plot_grid(",paste0("p",(8*i-8):(8*i-8+n_file),collapse=","),",align='V',ncol=1,rel_heights = c(0.1,0.1,0.1,0.1,0.1,0.1,0.1,0.1))")
      }
      
      commands3 <- paste0("print(g",i,")")
      commands4 <- "dev.off()"
      eval(parse(text=commands1))
      eval(parse(text=commands2))
      eval(parse(text=commands3))
      eval(parse(text=commands4))
    }
  
    return( eval(parse(text=paste0("list(n_pdf,",paste0("g",1:n_pdf,collapse=","),")"))))
}

combine_sig_nmf <- function(input_folder,cancertype){
  
  for (i in 1:length(cancertype)){
  tryCatch({
    type <- cancertype[i]
    print(paste0("load ",i," th type : ",type))
    
    sig_file <- dir(paste0(input_folder,type,"/"))[grep("sig",dir(paste0(input_folder,type,"/")))]
    load(paste0(input_folder,type,"/",sig_file))
  
    colnames(sig) <- paste0(type,"_sig",1:ncol(sig))
    if (i==1) combine_sig <- sig else
       combine_sig <- cbind(combine_sig,sig)
  },error=function(e) print("Fail load this type"))
  }  
   save(combine_sig,file=paste0(input_folder,"combine_sig_",Sys.Date(),".RData"))
   return(combine_sig)
}

hc_cluster_test <- function(data,methods,distance,min = 2,max = 10){
  library(NbClust)
  
  getmode <- function(v) {
    uniqv <- unique(v)
    uniqv[which.max(tabulate(match(v, uniqv)))]
  }
  
  tabla = as.data.frame(matrix(ncol = length(distance), nrow = length(methods)))
    names(tabla) = distance
  
  for (j in 2:length(distance))
      for(i in 1:length(methods)){
        tryCatch({
        nb = NbClust(data,distance = distance[j],
                     min.nc = min, max.nc = max, 
                     method = "complete", index =methods[i])
        tabla[i,j] = nb$Best.nc[1]
        tabla[i,1] = methods[i]
      },error=function(e) print("error"))
    } 
  
  tabla <- rbind(tabla,c("Most_common",apply(tabla[,2:5],2,getmode)))
  
  return(tabla)
}

## Step 8: run Hierarchical clustering and plot (save hc_heatmap, consensus sig and plot to output folder)
hc_consensus <- function(combine_sig,cluster,ccfMatrix,output,distance="euclidean")  {
  library("RColorBrewer")
  library("pheatmap")
  #library(factoextra)
  
  upper = quantile(combine_sig,0.95)
  breaksList = seq(0, upper, by = 0.01)
  col <- colorRampPalette(rev(brewer.pal(n = 6, name = "RdYlBu")))(length(breaksList))
    
  ## set `cutree_cols` based on suggested cluster number 
  out <- pheatmap(combine_sig, cutree_cols = cluster, fontsize_col = 5,fontsize_row = 0.4,color = col, breaks = breaksList,clustering_distance_cols=distance, cluster_rows=F,filename=paste0(output,"/",distance,"_",cluster,"_hc_heatmap.pdf"),clustering_method = "ward.D2")
 
  sig_label <- as.data.frame(cutree(out$tree_col,k=cluster)) %>%
      set_colnames("cluster") %>%
      mutate(sig=rownames(.))
    
  ## Compute consensu signatures for each cluster
  consensus_sig <- as.data.frame(t(combine_sig)) %>%
      mutate(sig=rownames(.)) %>%
      left_join(.,sig_label,by="sig") %>%
      group_by(cluster) %>%
      mutate(sig=NULL) %>%
      summarise_all(mean) 
  
  save(consensus_sig,file=paste0(output,"/",distance,"_",cluster,"_consensus_sig.RData"))
  consensus_sig <- apply(t(consensus_sig[,2:101]),2,as.numeric)
  
  p1 <- plot_grid(sig_plot(consensus_sig))
  save_plot(paste0(output,"/",distance,"_",cluster,"_consensus_sig.pdf"),p1,base_asp = cluster)
  
  if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
  if (!requireNamespace("YAPSA",quietly = TRUE)) BiocManager::install("YAPSA")
  library(YAPSA)
  
  load(paste0(HC_folder,dir(HC_folder)[grep("samples",dir(HC_folder))]))
  ccfMatrix[is.na(ccfMatrix)] = 0
  idx_random <- sample(1:nrow(ccfMatrix))
  ccfMatrix <- ccfMatrix[idx_random,]
  samplelist <- post_summary$samplename[idx_random]
  
  write.csv(samplelist,file=paste0(HC_folder,distance,"_",cluster,"_sample_list_random.csv"))
  
  exposure <- YAPSA::LCD(ccfMatrix,consensus_sig) 
  table(rowSums(ccfMatrix)>0)
  exposure <- as.data.frame(t(exposure)) %>%
    set_colnames(paste0("sig_",1:ncol(.)))
  save(exposure,file=paste0(output,"/",distance,"_",cluster,"_lcd_exposure.RData"))
  
  return(consensus_sig)
}


```


# Main

```{r}
## pre-step: set directory and Initialization
  library(readxl)
  
  rm(ls())
  
  ## Initialization - TCGA
  setwd("~/Downloads/ED_project/Signature_extract/test_TCGA")
  
  Ccube_folder <- "1.Ccube/TCGA_0106/"
  Postsummary_folder <- "2.post_summary/"
  Matrix_folder <- paste0("3.ccfMatrix/")
  rank_estimate_folder <- paste0(Matrix_folder,"rank_estimate/")
  rank_result_folder <-  paste0("4.rank_result/")
  nmf_folder <- "5.NMF_result/"
  HC_folder <- "6.HC_consensus/"

  cancertype_path <- "1.Ccube/TCGA_cancertype.csv"
 
  ## Initialization - ICGC
  setwd("~/Downloads/ED_project/Signature_extract/test_TCGA")
  
  Ccube_folder <- "1.Ccube/ICGC/"
  Postsummary_folder <- "2.post_summary/"
  Matrix_folder <- paste0("3.ccfMatrix/")
  rank_estimate_folder <- paste0(Matrix_folder,"rank_estimate/")
  rank_result_folder <-  paste0("4.rank_result/")
  nmf_folder <- "5.NMF_result/"
  HC_folder <- "6.HC_consensus/"

  
  cancertype_path <- "1.Ccube/ICGC_post_summary_type_2778_191107.csv"
  
  ## mkdir
  
  
  multi.dir.create <- function(list){
    for (i in list) {if (!dir.exists(i)) dir.create(i,recursive = T)}
  }
 
  multi.dir.create(c(Ccube_folder,Postsummary_folder,Matrix_folder,rank_result_folder,nmf_folder,HC_folder,lcd_folder,association_folder,rank_estimate_folder))
     
#####################################################################  
## Step1: run Ccube @ccube_test_v2.R | results already saved in test folder
  
#####################################################################  
## Step2: build post_summary
  
  post_summary <- Build_post_summary(input=Ccube_folder,output=Postsummary_folder,mergeType = T, typefile = cancertype_path)
  #post_summary <- load("2.post_summary/post_summary_2778_2020-01-07.RData")
      ## summary plot
      xx <- post_summary %>% group_by(cancertype) %>% summarize(n=n())
      
      library(ggplot2)
      ggplot(data=xx,aes(x=cancertype,y=n)) +
        geom_bar(position = 'dodge', stat='identity') +
        geom_text(aes(label=n), position=position_dodge(width=0.9), vjust=-0.25)
      ggsave(paste0(Postsummary_folder,"type_summary.pdf"),width = 50, height = 15, units = "cm")
      ## filter types with >30 samples
      types <- xx %>% filter(n>30) %>% select(cancertype) 
      post_summary <- subset(post_summary,cancertype %in% types$cancertype)
      
  save(post_summary,file=paste0(HC_folder,nrow(post_summary),"_analyased_samples.RData"))
  #save(post_summary,file='/home/xinyu/Downloads/ED_project/Signature_extract/results/TCGA/6.HC_consensus/7740_analyased_samples.RData')
######################################################################################  
## Step 3: Construct ccf matrix for samples in the post_summary by types and for all and save the result to Matrix_folder
  TypeMatrixBuild(post_summary = post_summary, input_folder=Ccube_folder,output=Matrix_folder,ccfupper = 1)  
  
###############################################################################  
#============================= go to python ==================================#
## Step 4: Rank estimate ==> output rank_summary ("4.rank_summary.xlsx") 
  
  # modify python script manually
  # -test.py: set path = `rank_estimate_folder`, set output = `rank_result_folder`
  # -estimate_plot.py : set output = `rank_result_folder`, set export_csv = output
  
  # run python script: 1.test.py 2.estimate_plot.py | results already saved in `rank_result_folder`
  
  # Generate rank estimate plots (go back to R) and "4.rank_summary_test.csv" with suggested rss rank
  
  rank_file_blank <- "4.rank_summary_test_blank.csv"
  
  rank_estimate_plot(outputFolder = rank_result_folder,rankfilepath = rank_file_blank)
  
  # * fill manually $rank in "4.rank_summary_test_blank.csv" and change name to "4.rank_summary_test.csv"
######################################################  
## Step 5: NMF for each type based on specified ranks
  
  # Notice: must fill the rank_summary sheet first to specify ranks
  ## make sure first column(cancertype) and 2nd column(rank)
  rank_file <- "4.rank_summary_test.csv"
  rank_summary <- read.csv(file=rank_file)[,-1]
  
  # run NMF for each cancer type 
  nmf_sig_all_plot(input_folder=Matrix_folder,output=nmf_folder,rank_summary=rank_summary)
  
########################################################  
## Step 6: HC + LCD
## Combine signatures extracted from differernt cancer types, run Hierarchical cluster and plot (save hc_heatmap, consensus sig and plot to output folder)
  
  ## combine signatures
  combine_sig <- combine_sig_nmf(input_folder= nmf_folder,cancertype=rank_summary$cancertype) %>%
    apply(2,function(x) x/sum(x)) 
  
  ## test the number of cluster
  lista.methods = c("kl","ch","hartigan",
                      "cindex","db","silhouette","ratkowsky","ball",
                       "ptbiserial","gap", "frey", "mcclain",  "gamma", "gplus", "tau", "dunn", 
                       "sdindex", "sdbw") # "hubert","dindex"
  lista.distance = c("metodo","euclidean", "maximum", "manhattan", "canberra")
    
  hc_cluster_test(data=combine_sig,methods=lista.methods,distance=lista.distance,max=10,min=2) 
  
  #=================== specify cluster numbers ==============================#
  ## run hierarchical cluster based on specified cluster and use LCD to get the exposure (save in folder `6.HC_consensus`)
  ## check all the results in folder `6.HC_consensus`
  ## Construct ccf matrix for all samples
  ccfMat <- CountMatBuild(samplelist=post_summary$samplename,upper=1,input_folder=Ccube_folder)
  save(ccfMat,file=paste0(Matrix_folder,"CCFMatrix_overall.RData"))
  #load(file=paste0(Matrix_folder,"CCFMatrix_overall.RData"))
  
  consensus_sig_eu <- hc_consensus(combine_sig,cluster=3,ccfMatrix=ccfMat[[3]],output=HC_folder,distance="euclidean")
  consensus_sig_correlation <- hc_consensus(combine_sig,cluster=3,ccfMatrix=ccfMat[[3]],output=HC_folder,distance="correlation")

```











